Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-03-16 Ben_Manuscripts/thesis/intro.tex
Intro outline ready for a lookX

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.
    \item The diversity in size and chemical topology of contaminants makes the
    design of membrane processes meant to separate these components complex.
    \item To deal with complex aqueous streams, and to prevent excessive membrane
    fouling, multiple types of membranes are often used in series, removing 
    larger particles first.~\cite{goosen_fouling_2005}
  \end{itemize}

@@

  Membranes are generally designed to maximize two key performance metrics: 
  permeability and selectivity.
  \begin{itemize}
    \item Permeability is used to quantify the rate at which solvent and solutes
    pass through the membrane.
    \item Selectivity quantifies the relative rates of passage of two substances.
    \item A related quantity to selectivity, rejection, quantifies the fraction
    of undesired solute that does not pass through the membrane.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.~\cite{werber_materials_2016}
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi et al. calculated that tripling membrane permeability
    could reduce energy consumption up to 15\% in a seawater RO plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment.
    \item Addressing these problems with highly selective separations may lead to 
    more gains.
    %MRS: math (or citations) for this statement?
 \end{itemize}

@@

  %BJC2: this section needs meat and citations

  Selective separations are useful in a wide range of applications.

  \textit{Desalination}:
  Creating potable water from seawater or brackish water is of paramount importance
  in water-scarce areas. 
  \begin{itemize}
  	\item Compared with thermal distillation techniques, RO has been shown to 
  	be more environmently friendly and economical due to its lower energy
  	requirements.\cite{morton_environmental_1997}
 	\item In general, RO has a smaller footprint % BJC: need citation
  	\item Thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.\cite{bhojwani_technology_2019}
  \end{itemize}

  \textit{Organic Micropollutants}:
  Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which can have adverse effects on human health even at low 
  concentrations.\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceuticals, personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply.
    %BJC2: There is a review I need to wade through to bulk this up.
    \item Is there any infrastructure built to combat this?
  \end{itemize}

  \textit{Recovery of Valuable Dissolved Species}: We can use highly selective
  membranes in order to recover potentially valuable dissolved species from 
  complex waste streams.

  Municipal waste-waters are rich in carbon, nitrogen and phosphorus-containing 
  compounds. % i.e. poop
  \begin{itemize}  
    \item The recovery of such products, which can be achieved using a 
    selective membrane, has numerous potential uses.\cite{sales_resource_2015}
    \item For example, nitrogen and phosphorus recovery can help sustain fertilizer 
    production which will help meet global food demand as population continues
    to increase.\cite{xie_membrane-based_2016}
  \end{itemize}

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater\cite{werber_materials_2016}. 
  \begin{itemize}
    \item For example, flowback water, produced during hydraulic fracturing of
    shale formations consists of relatively high concentrations of salts, metals,
    and soluble organic compounds. 
    \item The majority of this water is disposed through deep well injection, 
    however there is a growing public concern about its management which has 
    prompted the use of separation technologies such as RO and NF in order to 
    reduce the volume of contaminated water.~\cite{gregory_water_2011}
    \item Some of the dissolved organic compounds in flowback water, such as acetate, 
    are potentially valuable and can be recovered with highly selective membranes
    \cite{dischinger_application_2017}.
  \end{itemize}

  \textit{Breathable barriers}:

  Finally, there is a great deal of military interest in creating breathable 
  barriers which selectively allow the passage of water in both directions but
  blocks out harmful contaminants.
  \begin{itemize}
    \item Need to allow sweating
    \item Can even put catalytic groups in the pores to break down contaminants
  \end{itemize}

  \section{Competing Membrane Technologies}
  %BJC2: maybe better to title this section "Historical Development of Membrane Technology"

@@

  Nanofiltration membranes have pores on the order of 1 nm in size. Larger pore sizes 
  allow lower operating pressures (3 -- 20 bar) which lowers energy consumption.~\cite{van_der_bruggen_review_2003}
  An ideal NF membrane should have densely packed, uniform-sized and non-tortuous pores.
  This combination has been practically difficult to realize.

  The most widely used methods of NF membrane fabrication are interfacial polymerization
  and phase inversion.
  \begin{itemize}
    \item By modifying the monomer materials, one can control the structural morphology
    produced by interfacial polymerization to create pores larger than those in RO
    membranes.~\cite{seman_nanofiltration_2010}
    \item During phase inversion, one submerges a polymer, dissolved in a solvent, in
    a non-solvent. A solid, porous polymer membrane is all that remains once all
    solvent has been removed by non-solvent exchange.~\cite{smolders_microstructures_1992}
    \item Although commercially dominant, both of these techniques suffer from
    tortuous pores and polydisperse pore sizes which results in far from ideal 
    selectivities.
  \end{itemize}

  Track-etching is another technique which attempts to eliminate polydispersity and tortuosity.
  During track-etching, a polymer film is bombarded with charged particles, then chemically 
  etched to create pores.~\cite{apel_track_2001}
  \begin{itemize}
	\item Although the pores are uniform and non-tortuous, benefitting selectivity, the 
	membranes have an extremely low pore density subsequently low permeability. 
  \end{itemize}

  %BJC2: maybe one more paragraph here with applications

@@

  %BJC2: Each tecnhology below will likely have its own heading and be split into 2 or more 
  % paragraphs.

  Ultrathin-film graphene and graphene oxide membranes are an active area of research because
  they are atomically thick materials and therefore offer potential for extremely high
  permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}
    \item The tightly packed aromatic network of $sp^2$ hybridized carbon atoms that 
    make up graphene sheets precludes the passage of even the smallest monoatomic species, helium.~\cite{bunch_impermeable_2008}
    \item Controllably-sized pores have been created in pristine graphene by plasma etching
    and by charged particle bombardment.~\cite{surwade_water_2015,russo_atom-by-atom_2012}
    \item Graphene oxide (GO), a highly oxidized modification of graphene, consist of epoxy,
    hydroxyl and carboxylate functional groups.~\cite{hegab_graphene_2015}
    \item Many groups study GO membranes because they have improved antibacterial properties 
    over graphene, leading to less membrane fouling, and because their functionality can be
    modified in order to achieve solute-specific separations.~\cite{liu_antibacterial_2011,he_bioinspired_2013}
    \item Despite their promise, scalable synthesis of monolayer graphene and GO membranes without
    introducing microscopic, performance degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
    \item Multilayered GO membranes have shown considerably more progress since their 
    synthesis procedures are low cost and scalable.~\cite{zhang_size-controlled_2009}
    \item GO sheets are particularly easy to make into multilayered membranes since they 
    stack on top of each other, held together by hydrogen bonds.~\cite{homaeigohar_graphene_2017}
    \item The potential of graphene and GO membrane certainly merits their continued exploration. Many
    groups are still working in order to work on better control and standardization of their synthesis,
    improving their designability for specific separations, and further improving their scalability.~\cite{wei_multilayered_2018}
  \end{itemize}

  Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011}
  \begin{itemize}
    \item CNTs are an allotrope of carbon consisting of tubular graphitic layers.~\cite{iijima_helical_1991}
    \item One can control CNT diameters down to 0.5 nm in size which has inspired the
    creation of CNT membranes for reverse osmosis applications.~\cite{sun_creating_2000}
    \item Molecular dynamics simulations have predicted and experiments have confirmed that
    CNTs are extremely fast transporters of water.~\cite{hummer_water_2001,joseph_why_2008}
    \item MD simulations by Kalra et al. predicted water transport at rates comparable
    to the transmembrane protein aquaporin-1.~\cite{kalra_osmotic_2003}
    \item Experiments have exhibited water transport rates orders of magnitude higher than 
    predictions by continuum modeling.~\cite{holt_fast_2006}
    \item This enhancement has been explained as a result of nearly friction-less single-file water
    flow through the hydrophobic CNT interiors.~\cite{kalra_osmotic_2003}
    \item Functionalization of the CNT pore entrances can potentially lead to highly selective separations.~\cite{fornasiero_ion_2008}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them,
    however facile incorporation at scale remains a challenge yet to be overcome.~\cite{kim_polysulfone_2007}
  \end{itemize}

  Zeolite membranes offer the potential for permeabilities comparable to ultrafiltration, 
  with selectivities as good as NF and RO.~\cite{hoek_nanotechnology-based_2014}
  \begin{itemize}  
    \item Zeolites have highly uniform nm-sized crystalline structures with cage-like
    cavities that allow movement and trapping of small solutes. 
    \item The crystalline frameworks are typically formed by networks of silicon 
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
    \item One can replace the silicon and aluminum atoms via ion exchange in order
    to control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}
    \item Zeolite-coated ceramic membranes have shown the most promise for RO applications
    because of their mechanical stability under high applied pressure and chemical
    stability for withstanding disinfectants.~\cite{pendergast_review_2011}
    \item A number of studies have tested the permeability and sodium salt rejection
    of various zeolite membranes, however their high permeabilities typcially come
    at the cost of low selectivities and vice versa.~\cite{li_desalination_2004,duke_seawater_2009}
    \item Zeolites are also prone to defects in the crystalline structure which 
    would require substantial modification to the traditional synthesis procedures
    in order to fix.~\cite{kumakiri_application_2000}
  \end{itemize}

  Finally, metal organic frameworks (MOFs) have been recently introduced to the
  field of aqueous separations.
  \begin{itemize}
    \item MOFs are highly porous and flexible crystalline structures made of metal
    clusters coordinated to organic ligands.~\cite{furukawa_chemistry_2013}
    \item There is a much wider breadth of research into MOFs as gas separation membranes
    because they tend to be unstable in water.~\cite{feng_construction_2013}
    \item But a new generation of water-stable MOFs is making it possible to use them
    in aqueous separations.~\cite{wang_ultrastable_2015}
    \item Purification of both aqueous and gaseous streams using MOFs are 
    typically achieved by selective adsorption or catalytic degradation.~\cite{li_metalorganic_2012}    
    \item MOFs have been applied towards the removal of a number of common water 
    contaminants including dyes, pesticides, endocrine disruptors, and pharmaceuticals.
    \cite{zhang_fabrication_2015,seo_adsorptive_2015,li_metalorganic_2015,hasan_adsorptive_2012}
    \item Although aqueous MOF separations are theorized to surpass the performance 
    of conventional membrane separations, it is still a budding field. The stability and
    reusability of the material must be guaranteed before they can be implemented
    at scale.~\cite{dias_towards_2015}
  \end{itemize}

@@

  Amphiphilic molecules that self-assemble to form ordered nanostructures have shown
  potential as nanoporous membranes.
  \begin{itemize}
    \item Block copolymers
    \item Some other amphiphilic self-assembled thing
    \item Liquid crystals as the guy we are studying because we have lots of 
    experimental data.
  \end{itemize}

  Under the right conditions, the shape and amphiphilic character of the monomers in 
  Figure TBD drives their self assembly into ordered nanostructures.
  \begin{itemize}
    \item Monomer 1 can form the inverted hexagonal phase (H\textsubscript{II})
    \item Monomer 2 will form the bicontinuous cubic phase (Q\textsubscript{I})
    \item The tails have vinyl groups which can be cross-linked for mechanical strength.
  \end{itemize}

  %BJC: figure here showing each monomer and the phase they form

  In this work we attempt to understand how one can design LLC membranes as 
  alternative nanostructured materials for highly selective aqueous separations.

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item The inhomogeneity of the pores suggests that the effective pore size may depend
    on the solute. Both the size, shape and chemical functionality will influence
    how a solute partitions within the membrane. %BJC2: need to rework this sentence.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

  Unfortunately, it is a difficult task to align the self-assembled hexagonal
  mesophases into continuous pores that traverse the thickness of the membrane.
  \begin{itemize}
    \item Although the membranes have shown high experimental selectivity, their 
    flux has been very low due to misalignment of the hexagonal mesophases.
  \end{itemize}

%  Recently, Feng et al. has come up with two methods which can be used to 
%  macroscopically align the hexagonal mesophases. 
  Considerable recent efforts have made progress towards the macroscopic
  alignment of the hexagonal mesophases.
  \begin{itemize}
    \item Feng et al. leverage the magnetic anisotropy of the columnar \
    mesophases in order to control their alignment with a magnetic field.~\cite{feng_scalable_2014}
    \item Feng et al. also used an approach called soft confinement which
    takes advantage of the hexagonal columns preference towards anchoring
    perpendicular to either a PDMS or glass substrate.~\cite{feng_thin_2016}
    \item Finally, Feng et al. designed a third technique which uses a structure
    directing molecule in order to template the assembly of a fatty acid 
    molecule into ordered columnar phases.~\cite{feng_highly_2017}
    \item Aligned hexagonal phase membranes exhibited an 85 fold increase in 
    ionic conductivity.~\cite{feng_scalable_2014}
  \end{itemize}

@@

  The bicontinuous cubic, or Q\textsubscript{I} phase, is a class 
  of nanostructured phases characterized by a network of 3 dimensionally
  interconnected pores.
  \begin{itemize}
    \item Aside from its tortuous pore architecture, it shares the 
    uniform size and complex topological features of the H\textsubscript{II} phase
    which lends itself to highly selective separations.
    \item Although water and solutes must follow a longer path in order to pass
    through Q\textsubscript{I} membranes, they are more practical to synthesize
    because the mesophases do not require alignment. 
  \end{itemize}

@@

  The next four chapters offer answers to each of the above questions and are adapted from 
  various first author manuscripts at various stages in the publication process as described
  below. The introduction to each adaptation is condensed in order to avoid redundancy and 
  all other sections are left intact. The supporting information documents for each manuscript are
  given in the appendices. \\

>>>>

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers.~\cite{goosen_fouling_2005}
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.~\cite{werber_materials_2016}
    \item The diversity in size and chemical topology of contaminants makes the
    design of membrane processes meant to separate these components complex.
    \item To deal with complex aqueous streams, and to prevent excessive membrane
    fouling, multiple types of membranes are often used in series, removing 
    larger particles first.~\cite{suk_membranebased_2006}
  \end{itemize}

@@

  Membranes are generally designed to optimize two key performance metrics: 
  permeability and selectivity.~\cite{geise_water_2011}
  %BJC2: Not sure the right place to put citation since its all described in the above citation. I could add more, but seems pointless.
  \begin{itemize}
    \item Permeability is used to quantify the rate at which solvent and solutes
    pass through the membrane.
    \item Selectivity quantifies the relative rates of passage of two substances.
    \item A related quantity to selectivity, rejection, quantifies the fraction
    of undesired solute that does not pass through the membrane.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.~\cite{werber_materials_2016}
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi \textit{et al.} calculated that a theoretical tripling of membrane permeability
    would only reduce energy consumption by up to 15\% in a seawater RO plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.~\cite{singh_production_2009}
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment.~\cite{shannon_science_2009,fritzmann_state---art_2007}
%    \item The well-known permeability-selectivity tradeoff, common to dense polymer 
%    membranes, directly correlates an increase in water permeability with an 
%    increase in solute permeability.~\cite{geise_water_2011}
    \item Addressing these problems with highly selective separations may lead to 
    more gains.~\cite{deshmukh_desalination_2015}
 \end{itemize}

@@

  Selective separations are useful in a wide range of applications. They can be used
  to separate components in both gaseous and aqueous phases.~\cite{baker_gas_2014}. We 
  will focus this section on aqueous separations since they are the target process
  for the following studies.

  \textit{Desalination}: Creating potable water from seawater or brackish water is of 
  paramount importance in water-scarce areas. 
  \begin{itemize}
    \item It is estimated that by 2025, 1.8 billion people may live in water-stressed
    regions.~\cite{navarro-ortega_managing_2015} %There's a water stress index in fritzmann_2007
    \item Membrane deslination via RO and thermal muli-stage flash distillation are 
    both widely used processes for creating potable water in these regions.~\cite{fritzmann_state---art_2007}
  	\item Compared with thermal distillation techniques, RO has been shown to 
  	be more environmentally friendly and economical due to its lower energy
  	requirements.\cite{morton_environmental_1997}
% 	\item In general, RO has a smaller footprint % BJC: need citation
  	\item However, thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.\cite{bhojwani_technology_2019}
  \end{itemize}

  \textit{Organic Micropollutants}: Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which can have adverse effects on human health even at low 
  concentrations.\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceuticals, personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply.~\cite{barbosa_occurrence_2016}
    \item Sources of micropollutants incude industrial wastewater, 
    agricultural runoff, landfill leachates and domestic effluents.~\cite{mompelat_occurrence_2009}
    %BJC2: There is a review I need to wade through to bulk this up.
    \item Unfortunately, most waste water treatment plants are not 
    designed for complete removal of these contaminants.~\cite{tijani_review_2013}
    \item There is a need for a new generation of membranes that can better
    target specific molecules.
  \end{itemize}

  \textit{Recovery of Valuable Dissolved Species}: In addition to removing contaminants 
  to clean waste water, we can also use highly selective membranes in order to recover
  potentially valuable dissolved species from complex municipal and industrial waste
  streams.~\cite{guest_new_2009,daigger_evolving_2009}

  Municipal waste-waters are rich in carbon, nitrogen and phosphorus-containing 
  compounds.~\cite{romero_raw_2013}
  \begin{itemize}  
    \item The recovery of such products, which can be achieved using a 
    selective membrane, has numerous potential uses.\cite{sales_resource_2015}
    \item For example, nitrogen and phosphorus recovery can help sustain fertilizer 
    production which will help meet global food demand as population continues
    to increase.\cite{xie_membrane-based_2016}
  \end{itemize}

  Industrial waste-waters are often quite complex with up to six times more 
  total dissolved solids than seawater.\cite{werber_materials_2016} 
  \begin{itemize}
    \item For example, flowback water, produced during hydraulic fracturing of
    shale formations consists of relatively high concentrations of salts, metals,
    and soluble organic compounds.~\cite{orem_organic_2014}
    \item The majority of this water is disposed through deep well injection, 
    however there is a growing public concern about its management which has 
    prompted the use of separation technologies such as RO and NF in order to 
    reduce the volume of contaminated water.~\cite{gregory_water_2011}
    \item Some of the dissolved organic compounds in flowback water, such as acetate, 
    are potentially valuable and can be recovered with highly selective membranes
    \cite{dischinger_application_2017}.
  \end{itemize}

%BJC: this is more of a gas separation
%  \textit{Breathable barriers}:
%  
%  Finally, there is a great deal of military interest in creating breathable 
%  barriers which selectively allow the passage of water in both directions but
%  blocks out harmful contaminants.
%  \begin{itemize}
%    \item Need to allow sweating
%    \item Can even put catalytic groups in the pores to break down contaminants
%  \end{itemize}

  \section{Historical Development of Membrane Technology}
  %BJC2: maybe better to title this section "Historical Development of Membrane Technology"

@@

  Since their introduction in the late 1980s, NF membranes have been applied towards
  an increasing number of separations since they bridge the gap in capabilities
  between reverse osmosis and ultrafiltration.~\cite{eriksson_nanofiltration_1988}
  \begin{itemize}
    \item Nanofiltration membranes have pores on the order of 1 nm in size. 
    \item Larger pore sizes allow lower operating pressures (3 -- 20 bar) which lowers 
    energy consumption.~\cite{van_der_bruggen_review_2003}
    \item An ideal NF membrane should have densely packed, uniform-sized and non-tortuous pores.~\cite{jackson_nanoporous_2010}
    This combination has been practically difficult to realize.
  \end{itemize}

  The most widely used methods of NF membrane fabrication are interfacial polymerization
  and phase inversion.
  \begin{itemize}
    \item NF membranes created by interfacial polymerization are made similarly to
    RO membranes. By modifying the monomer materials, one can control the structural morphology
    in order to create pores larger than those in RO membranes.~\cite{seman_nanofiltration_2010}
    \item During phase inversion, one submerges a polymer, dissolved in a solvent, in
    a non-solvent. A solid, porous polymer membrane is all that remains once all
    solvent has been removed by non-solvent exchange.~\cite{smolders_microstructures_1992}
    \item Although commercially dominant, both of these techniques suffer from
    tortuous pores and polydisperse pore sizes which results in far from ideal 
    selectivities.
  \end{itemize}

  Track-etching is another NF fabrication technique which attempts to eliminate 
  polydispersity and tortuosity. During track-etching, a polymer film is bombarded with 
  charged particles, then chemically etched to create pores.~\cite{apel_track_2001}
  \begin{itemize}
	\item Although the pores are uniform and non-tortuous, benefiting selectivity, the 
	membranes have an extremely low pore density and subsequently low permeability. 
  \end{itemize}

@@

  %BJC2: Each tecnhology below will likely have its own heading and be split into 2 or more 
  % paragraphs.
  \textit{Graphene and graphene oxide} Ultrathin-film graphene and graphene oxide membranes
  are an active area of research because they are atomically thick materials and therefore 
  offer potential for extremely high permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}
    \item The tightly packed aromatic network of $sp^2$ hybridized carbon atoms that 
    make up graphene sheets precludes the passage of even the smallest monoatomic species, helium.~\cite{bunch_impermeable_2008}
    \item Controllably-sized pores have been created in pristine graphene by plasma etching
    and by charged particle bombardment.~\cite{surwade_water_2015,russo_atom-by-atom_2012}
    \item Graphene oxide (GO), a highly oxidized modification of graphene, consist of epoxy,
    hydroxyl and carboxylate functional groups.~\cite{hegab_graphene_2015}
    \item Many groups study GO membranes because they have improved antibacterial properties 
    over graphene, leading to less membrane fouling, and because their functionality can be
    modified in order to achieve solute-specific separations.~\cite{liu_antibacterial_2011,he_bioinspired_2013}
    \item Despite their promise, scalable synthesis of monolayer graphene and GO membranes without
    introducing microscopic, performance degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
    \item Multilayered GO membranes have shown considerably more progress since their 
    synthesis procedures are low cost and scalable.~\cite{zhang_size-controlled_2009}
    \item GO sheets are particularly easy to make into multilayered membranes since they 
    stack on top of each other, held together by hydrogen bonds.~\cite{homaeigohar_graphene_2017}
    \item The potential of graphene and GO membrane certainly merits their continued exploration. Many
    groups are still working in order to work on better control and standardization of their synthesis,
    improving their designability for specific separations, and further improving their scalability.~\cite{wei_multilayered_2018}
  \end{itemize}

  \textit{Carbon nanotubes}: Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011}
  \begin{itemize}
    \item CNTs are an allotrope of carbon consisting of tubular graphitic layers.~\cite{iijima_helical_1991}
    \item One can control CNT diameters down to 0.5 nm in size which has inspired the
    creation of CNT membranes for reverse osmosis applications.~\cite{sun_creating_2000}
    \item Molecular dynamics simulations have predicted and experiments have confirmed that
    CNTs are extremely fast transporters of water.~\cite{hummer_water_2001,joseph_why_2008}
    \item MD simulations by Kalra \textit{et al.} predicted water transport at rates comparable
    to the transmembrane protein aquaporin-1.~\cite{kalra_osmotic_2003}
    \item Experiments have exhibited water transport rates orders of magnitude higher than 
    predictions by continuum modeling.~\cite{holt_fast_2006}
    \item This enhancement has been explained as a result of nearly friction-less single-file water
    flow through the hydrophobic CNT interiors.~\cite{kalra_osmotic_2003}
    \item Functionalization of the CNT pore entrances can potentially lead to highly selective separations.~\cite{fornasiero_ion_2008}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them,
    however facile incorporation at scale remains a challenge yet to be overcome.~\cite{kim_polysulfone_2007}
  \end{itemize}

  \textit{Zeolites}: Zeolite membranes offer the potential for permeabilities comparable to ultrafiltration, 
  with selectivities as good as NF and RO.~\cite{hoek_nanotechnology-based_2014}
  \begin{itemize}  
    \item Zeolites have highly uniform nm-sized crystalline structures with cage-like
    cavities that allow movement and trapping of small solutes. 
    \item The crystalline frameworks are typically formed by networks of silicon 
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement.~\cite{auerbach_handbook_2003}
    \item One can replace the silicon and aluminum atoms via ion exchange in order
    to control the size of the cavities and hence its molecular-seiving properties.~\cite{li_novel_2007}
    \item Zeolite-coated ceramic membranes have shown the most promise for RO applications
    because of their mechanical stability under high applied pressure and chemical
    stability for withstanding disinfectants.~\cite{pendergast_review_2011}
    \item A number of studies have tested the permeability and sodium salt rejection
    of various zeolite membranes, however their high permeabilities typcially come
    at the cost of low selectivities and vice versa.~\cite{li_desalination_2004,duke_seawater_2009}
    \item Zeolites are also prone to defects in the crystalline structure which 
    would require substantial modification to the traditional synthesis procedures
    in order to fix.~\cite{kumakiri_application_2000}
  \end{itemize}

  \textit{Metal organic frameworks}: Finally, metal organic frameworks (MOFs) have been
  recently introduced to the field of aqueous separations.
  \begin{itemize}
    \item MOFs are highly porous and flexible crystalline structures made of metal
    clusters coordinated to organic ligands.~\cite{furukawa_chemistry_2013}
    \item There is a much wider breadth of research into MOFs as gas separation membranes
    because they tend to be unstable in water.~\cite{feng_construction_2013}
    \item But a new generation of water-stable MOFs is making it possible to use them
    in aqueous separations.~\cite{wang_ultrastable_2015}
    \item Purification of both aqueous and gaseous streams using MOFs are 
    typically achieved by selective adsorption or catalytic degradation.~\cite{li_metalorganic_2012}    
    \item MOFs have been applied towards the removal of a number of common water 
    contaminants including dyes, pesticides, endocrine disruptors, and pharmaceuticals.
    \cite{zhang_fabrication_2015,seo_adsorptive_2015,li_metalorganic_2015,hasan_adsorptive_2012}
    \item Although aqueous MOF separations are theorized to surpass the performance 
    of conventional membrane separations, it is still a budding field. The stability and
    reusability of the material must be guaranteed before they can be implemented
    at scale.~\cite{dias_towards_2015}
  \end{itemize}

@@

  Amphiphilic molecules that self-assemble to form ordered nanostructures have shown
  potential as nanoporous membranes.
  \begin{itemize}
    \item People are particularly interested in self-assembled materials because they
    offer the potential of uniform-sized pores and tunable chemical and physical 
    properties.~\cite{werber_materials_2016}
    \item There are two major classes of self-assembled amphiphilic materials: block 
    copolymers and liquid crystals.
  \end{itemize}

  Block copolymers contain more than one distinct repeating unit which facilitates 
  the formation of self-assembled structures based on differences in the subunit 
  amphiphilicity.~\cite{jackson_nanoporous_2010} 
  \begin{itemize}
%    \item The varying amphiphilic character of block copolymers, polymers with more than
%    one distinct repeating unit, allows them to form self-assembled structures.~\cite{jackson_nanoporous_2010} 
    \item Block copolymer membranes typically have pores on the order of 5-50 nm which
    makes them useful for UF applications, however there is progress being made towards
    smaller pores in the NF regime.~\cite{gu_tailoring_2015}
  \end{itemize}

  Under the right conditions, the shape and amphiphilic character of liquid 
  crystal monomers drives their self-assembling into ordered nanostructures with
  uniform-sized pores on the order of 1 nm in size.~\cite{gin_polymerized_2008}
  \begin{itemize}
    \item They have been used as membranes before.~\cite{gin_polymerized_2001}
    \item As shown in Figure TBD, monomer 1 will form the inverted hexagonal phase 
    (H\textsubscript{II})~\cite{smith_ordered_1997} while monomer 2 will form the 
    bicontinuous cubic phase (Q\textsubscript{I})~\cite{carter_glycerol-based_2012}
    \item The tails have vinyl groups which can be cross-linked for mechanical strength.
    \item In this work we attempt to understand how one can design LLC membranes as 
    alternative nanostructured materials for highly selective aqueous separations.
    \item We chose to study liquid crystals because they are already suitable as NF 
    membranes and there has been sufficient experimental characterization.~\cite{gin_polymerized_2001,feng_scalable_2014,feng_thin_2016}
  \end{itemize}

  %BJC: figure TBD here showing each monomer and the phase they form

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.~\cite{smith_ordered_1997}
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item The inhomogeneity of the pores suggests that the effective pore size may depend
    on the solute. Both the size, shape and chemical functionality will influence
    how a solute partitions within the membrane. %BJC2: need to rework this sentence.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

  Unfortunately, it is a difficult task to align the self-assembled hexagonal
  mesophases into continuous pores that traverse the thickness of the membrane.
  \begin{itemize}
    \item Although the membranes have shown high experimental selectivity, their 
    flux has been very low due to misalignment of the hexagonal mesophases.~\cite{zhou_supported_2005}
  \end{itemize}

%  Recently, Feng et al. has come up with two methods which can be used to 
%  macroscopically align the hexagonal mesophases. 
  Considerable recent efforts have made progress towards the macroscopic
  alignment of the hexagonal mesophases.
  \begin{itemize}
    \item Feng \textit{et al.} leveraged the magnetic anisotropy of the columnar
    mesophases in order to control their alignment with a magnetic field.~\cite{feng_scalable_2014}
    \item This study showed an 85-fold increase in ionic conductivity in the aligned
    versus unaligned state, demonstrating a huge increase in feasibility as an industrial membrane.
    \item Feng \textit{et al.} also used an approach called soft confinement which
    takes advantage of the hexagonal columns preference towards anchoring
    perpendicular to either a PDMS or glass substrate.~\cite{feng_thin_2016}
    \item Finally, Feng \textit{et al.} designed a third technique which uses a structure
    directing molecule in order to template the assembly of a fatty acid 
    molecule into ordered columnar phases.~\cite{feng_highly_2017}
%    \item Aligned hexagonal phase membranes exhibited an 85 fold increase in 
%    ionic conductivity.~\cite{feng_scalable_2014}
  \end{itemize}

@@

  The bicontinuous cubic, or Q\textsubscript{I} phase, is a class 
  of nanostructured phases characterized by a network of 3 dimensionally
  interconnected pores.~\cite{hyde_bicontinuous_1996}
  \begin{itemize}
    \item Aside from its tortuous pore architecture, it shares the 
    uniform size and complex topological features of the H\textsubscript{II} phase
    which lends itself to highly selective separations.
    \item Although water and solutes must follow a longer path in order to pass
    through Q\textsubscript{I} membranes, they are more practical to synthesize
    because the mesophases do not require alignment.~\cite{zhou_new_2007}
  \end{itemize}

@@

  The next four chapters offer answers to each of the above questions and are adapted from 
  my first author manuscripts at various stages in the publication process as described
  below. The introduction to each adaptation is condensed in order to avoid redundancy and 
  all other sections are left intact. The supporting information documents for each manuscript are
  given in the appendices. \\
